---
title: "P8105: Homework 3"
author: Alexander Melamed
output: github_document
---

```{r}
library(tidyverse)
library(ggplot2)
```


## Problem 1
Load instacart data
```{r}
library(p8105.datasets)
data("instacart")

```

**The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):**

The dataset `instacart` includes the following variables: `r names(instacart)`. This dataset has `r nrow(instacart)` rows and  `r ncol(instacart)` columns. Each row represents a single item in `r n_distinct(instacart$order_id)` orders. The varaiblase `order_ _dow` and `order_hour_of_day` indicates the day and time the order was placed.The variable `porduct_name` describes the item. The first three items in this dataset are: `r head(instacart$product_name, n=3)`.

**How many aisles are there, and which aisles are the most items ordered from?**

There are n_distinct `r n_distinct(instacart$aisle_id)` aisles, and the most items are ordered from aisle `r summarise(instacart, c = names(table(aisle_id))[which.max(table(aisle_id))])`.


**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

To make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered, I create a datafrom consisting of aisles and number of orders

```{r message=FALSE, warning=FALSE}
order_aisle = instacart %>% 
  group_by(aisle_id) %>% 
  summarise(n=n()) %>% 
  filter(n>10000) %>%
  arrange(n)%>%
  mutate(rank=rank(n))
```

Next I use ggplot to make a bar graph, with aisles arranged by increasing order number. I use the guide_axis option in scale_x_discrete to make the x-axis legible and correspond to the original aisle numbers. 

```{r}
ggplot(order_aisle, aes(x=factor(rank), y=n, )) +
  geom_bar(stat="identity")+
  theme_minimal()+
  labs(
    title = "Orders per aisle",
    x = "Aisle number",
    y = "Orders") +
  scale_x_discrete(
    labels=order_aisle$aisle_id,
    guide = guide_axis(n.dodge=2)
  )
```

**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
  instacart %>% 
  filter(aisle=="baking ingredients" | aisle=="dog food care" | aisle=="packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  arrange(-n) %>% 
  mutate(item_rank=rank(-n)) %>% 
  filter(item_rank<4) %>% 
  knitr::kable(digits = 1)

```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
  instacart %>% 
  filter(product_name== "Pink Lady Apples" | product_name=="Coffee Ice Cream") %>% 
  group_by(product_name, order_dow ) %>% 
  summarise(mean_hour_of_day = mean(order_hour_of_day)) %>% 
  mutate(day_of_week=
           ifelse(order_dow==0, "Sunday",
            ifelse(order_dow==1, "Monday",
              ifelse(order_dow==2, "Tueday",
               ifelse(order_dow==3, "Wedday",
                ifelse(order_dow==4, "Thursday",
                 ifelse(order_dow==5, "Friday", "Saturday"))))))) %>% 
  select(day_of_week, mean_hour_of_day) %>% 
  pivot_wider(
    names_from = day_of_week,
    values_from = mean_hour_of_day) %>% 
  knitr::kable(digits = 1) 
    
 
```
## Problem 2

**Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).**

```{r}
accel_data <- read_csv("data/accel_data.csv") %>%
  pivot_longer(
    activity.1:activity.1440,
    names_prefix="activity.",
    names_to="minute",
    values_to="activity_count"
  ) %>% 
  mutate(weekend=ifelse(day %in% c("Saturday", "Sunday"), TRUE, FALSE)) %>% 
  mutate(minute=as.numeric(minute))

```
The data set `accel_data` The  includes the following variables: `r names(accel_data)`. This dataset has `r nrow(accel_data)` rows and  `r ncol(accel_data)` columns. 

**Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?**

```{r}
accel_data %>%
  group_by(day_id,day,weekend) %>% 
  summarise(total_activity=
              sum(activity_count),
            ) %>% 
    knitr::kable(digits = 1)
  
```

Honestly no trends jump out at me from looking at this table. I tried a ridge plot to get more of a sense of how activity varied across day of the week:

```{r}
library(ggridges)
accel_data %>%
  group_by(day_id,day,weekend) %>% 
  summarise(total_activity=
              sum(activity_count),
            ) %>% 
    ggplot(aes(x=total_activity,y=day))+
  geom_density_ridges()

```
And a scatter with smooth to see how activity behaved over the study

```{r}
library(ggridges)
accel_data %>%
  group_by(day_id,day,weekend) %>% 
  summarise(total_activity=
              sum(activity_count),
            ) %>% 
    ggplot(aes(y=total_activity,x=day_id, color=day))+
  geom_point() 

```

All I might guess is that the person does a more constant amount of activity on tuesday to thrusday than on other days.

**Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph**


```{r}
library(plotly)
ggplot(accel_data, aes(x=minute, y=activity_count, color=day))+
  geom_point(alpha=0.05, size=1)+
  geom_smooth(se=FALSE)+
  scale_x_continuous(
    breaks = seq(from = 0, to = 1440, by = 120))+
  scale_y_continuous(trans = "log10")+
  scale_fill_viridis_c()  +
  theme(axis.text.x=element_text(angle=45,hjust=1))

```
We can see that on most day activity is minimal between minutes 0 and 240, and then increases. Activity drops off again after at approximately minute 1260. Saturday appears to be a lazier day. Minute-by-minute activity varies considerably. 
