---
title: "P8105: Homework 3"
author: Alexander Melamed
output: github_document
---

```{r}
library(tidyverse)
```


## Problem 1
Load instacart data
```{r}
library(p8105.datasets)
data("instacart")

```


The dataset `instacart` includes the following variables: `r names(instacart)`. This dataset has `r nrow(instacart)` rows and  `r ncol(instacart)` columns. Each row represents a single item in `r n_distinct(instacart$order_id)` orders. The varaiblase `order_ _dow` and `order_hour_of_day` indicates the day and time the order was placed.The variable `porduct_name` describes the item. The first three items in this dataset are: `r head(instacart$product_name, n=3)`.


There are n_distinct `r n_distinct(instacart$aisle_id)` aisles, and the most items are ordered from aisle `r summarise(instacart, c = names(table(aisle_id))[which.max(table(aisle_id))])`.


To make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered, I create a datafrom consisting of aisles and number of orders

```{r message=FALSE, warning=FALSE}
order_aisle = instacart %>% 
  group_by(aisle_id) %>% 
  summarise(n=n()) %>% 
  filter(n>10000) %>%
  arrange(n)%>%
  mutate(rank=rank(n))
```

Next I use ggplot to make a bar graph, with aisles arranged by increasing order number. I use the guide_axis option in scale_x_discrete to make the x-axis legible and correspond to the original aisle numbers. 

```{r}
ggplot(order_aisle, aes(x=factor(rank), y=n, )) +
  geom_bar(stat="identity")+
  theme_minimal()+
  labs(
    title = "Orders per aisle",
    x = "Aisle number",
    y = "Orders") +
  scale_x_discrete(
    labels=order_aisle$aisle_id,
    guide = guide_axis(n.dodge=2)
  )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
  instacart %>% 
  filter(aisle=="baking ingredients" | aisle=="dog food care" | aisle=="packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  arrange(-n) %>% 
  mutate(item_rank=rank(-n)) %>% 
  filter(item_rank<4) %>% 
  knitr::kable(digits = 1)

```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
  instacart %>% 
  filter(product_name== "Pink Lady Apples" | product_name=="Coffee Ice Cream") %>% 
  group_by(product_name, order_dow ) %>% 
  summarise(mean_hour_of_day = mean(order_hour_of_day)) %>% 
  mutate(day_of_week=
           ifelse(order_dow==0, "Sunday",
            ifelse(order_dow==1, "Monday",
              ifelse(order_dow==2, "Tueday",
               ifelse(order_dow==3, "Wedday",
                ifelse(order_dow==4, "Thursday",
                 ifelse(order_dow==5, "Friday", "Saturday"))))))) %>% 
  select(day_of_week, mean_hour_of_day) %>% 
  pivot_wider(
    names_from = day_of_week,
    values_from = mean_hour_of_day) %>% 
  knitr::kable(digits = 1) 
    
 
```

